Phonon polaritonics in two-dimensional materials

### 2019-03-18 paper.tex
minor obvious fixes

The third comment is that the considerations of this section can be extended to few-layer systems, such as hBN bilayers, which may still be considered 2D for polariton wavelengths large compared to the bilayer thickness. The considerations may also be extended to 1D polar materials, such as BN nanotubes. For bilayers, as was discussed in Ref.~\citenum{sohier2017breakdown}, one may describe the LO--TO splitting in terms of Eq.~\eqref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system. For a small number of layers $N$, this should enhance the LO--TO splitting by a factor of $N$. To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the "2.5D" Coulomb interaction \cite{thygesen2017calculating} in Eq.~\ref{eq:zerosofresponse}. In that case, for a film of thickness $d$, the interaction takes the form $V(q) = \frac{e^2}{\epsilon_0 dq^2}\left(1-\frac{2}{qd}e^{-\frac{qd}{2}}\sinh\left(\frac{qd}{2}\right) \right)$. This form interpolates between the 2D and 3D interactions in the limits $qd \rightarrow 0$ and $qd \rightarrow \infinity$. For sufficiently large thickness and finite-q, the 2.5D Coulomb interaction scales as $\frac{1}{dq^2}$. Noting the $q^2$ factor in Eq.~\ref{eq:zerosofresponse}, as well as the $N$-dependence of the quantity involving eigendisplacements, one finds an LO-TO splitting independent of thickness, as expected for bulk samples.

@@

In the rest of this section, we discuss the relation between the dispersion of an hBN monolayer versus the atomically-thick limit of a thin film of a material with hBN's bulk dielectric function. To aid this discussion, in Figure~\ref{fig:2}(c), we show the dispersion relation of thin films of bulk hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly \SI{0.33}{\nm}. For these plots, we take hBN to be cleaved such that the optical axis is perpendicular to the plane of the film. The components of the bulk permittivity perpendicular and parallel to the c-axis ($\epsilon_{\perp}$ and $\epsilon_{\parallel}$, respectively; indexed by $\alpha\in\{\perp,\parallel\}$ below) are then given by
\begin{equation}
	\epsilon_{\alpha}(\omega) = \epsilon_{\infty,\alpha}\Bigg(1 + \frac{\omega^2_{\mathrm{LO},\alpha}-\omega^2_{\mathrm{TO},\alpha}}{\omega^2_{\mathrm{TO},\alpha}-\omega^2} \Bigg),
	\label{eq:3deps}
\end{equation}
with $\epsilon_{\infty,\parallel} = \num{2.95}$, $\omega_{\mathrm{TO},\parallel} = \SI{760}{\per\cm}$, and $\omega_{\mathrm{LO},\parallel} = \SI{830}{\per\cm}$; and $\epsilon_{\infty,\perp} = \num{4.87}$, $\omega_{\mathrm{TO},\perp} = \SI{1360}{\per\cm}$, and $\omega_{\mathrm{LO},\perp} = \SI{1614}{\per\cm}$\cite{caldwell2014sub,dai2014tunable}.
Losses are ignored in this discussion altogether. 
In the range between $\omega_{\mathrm{TO}, \alpha}$ and $\omega_{\mathrm{LO}, \alpha}$, the corresponding component of the permittivity is negative, while the other component is positive. This hyperbolicity leads to a dispersion for hBN thin films that have multiple branches at a given frequency, as can be seen in Figure~\ref{fig:2}(c).  This trend persists even when the thickness of the bulk is taken down to a single layer, albeit being pushed to high wavevectors. This is in contrast with the true monolayer, in which there is only one LO phonon mode. Given that the phonon polariton of 2D is the LO phonon, there can be only one branch of the dispersion. The bulk dielectric function of Eq.~\eqref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity. For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode. However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of \SI{12}{\nm}, which is very short, the polariton frequency is only a few \si{\per\cm} above the TO frequency, within the damping linewidth of the TO frequency. %\textcolor{blue}{The bulk dielectric function of Eq.~\ref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity. For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode. However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Fig.~\ref{fig:2}(c), because even for a polariton wavelength of 12 nm, which is very short, the polariton frequency is only a few cm$^{-1}$ above the TO frequency, within the damping linewidth of the TO frequency.}

Another anomaly between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector. This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting. This turn-back point corresponds to a frequency of around \SI{1500}{\per\cm} and a phonon wavelength of less than \SI{1}{\nm}. Thus, this deviation is challenging to probe in typical optical experiments. However it means that for the monolayer, the range of accessible phonon polaritonic frequencies for optical applications is reduced compared to thick films. Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the one-atom-thick thin film suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit. This qualitative similarity is to be somewhat expected, as in many cases, a 2D layer can be considered as an very thin film of bulk, provided that the microscopic properties of the bulk and monolayer are the same. In fact, if we take an thin film of thickness $t$ of hBN of dielectric function given by Eq.~\eqref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk. It is: 
\begin{equation}
(\omega^{2 (\mathrm{bulk})}_{\mathrm{LO}} - \omega^{2 (\mathrm{bulk})}_{\mathrm{TO}})t = 4\omega^{(\mathrm{mono})}_{\mathrm{TO}}v^{(\mathrm{mono})}_g,
\end{equation}
where `bulk' denotes a property of the bulk dielectric function and `mono' denotes that of the monolayer conductivity.

>>>>

The third comment is that the considerations of this section can be extended to few-layer systems, such as hBN bilayers, which may still be considered 2D for polariton wavelengths large compared to the bilayer thickness. The considerations may also be extended to 1D polar materials, such as BN nanotubes. For bilayers, as was discussed in Ref.~\citenum{sohier2017breakdown}, one may describe the LO--TO splitting in terms of Eq.~\eqref{eq:zerosofresponse}, with the 2D Coulomb interaction, but with eigendisplacements appropriate to the few layer system. For a small number of layers $N$, this should enhance the LO--TO splitting by a factor of $N$. 
To go beyond an effectively 2D description of a monolayer, and make the transition from monolayer to bulk, we may make use of the quasi-2.5D Coulomb interaction~\cite{thygesen2017calculating} in Eq.~\eqref{eq:zerosofresponse}. 
In that case, for a film of effective thickness $d$, the interaction is approximately $V_{\text{Q2.5D}}(q) = \frac{e^2}{\epsilon_0 dq^2}\big(1-\frac{2}{qd}e^{-\frac{qd}{2}}\sinh\frac{qd}{2} \big)$. This form interpolates between the 2D and 3D interactions in the limits $qd \rightarrow 0$ and $qd \rightarrow \infty$. For sufficiently large thickness and finite-$q$, the 2.5D Coulomb interaction scales as $d^{-1}q^{-2}$. Noting the $q^2$ factor in Eq.~\eqref{eq:zerosofresponse}, as well as the $N$-dependence of the quantity involving eigendisplacements, one finds an LO--TO splitting independent of thickness, as expected for bulk samples.

@@

In the rest of this section, we discuss the relation between the dispersion of an hBN monolayer versus the atomically-thick limit of a thin film of a material with hBN's bulk dielectric function. To aid this discussion, in Figure~\ref{fig:2}(c), we show the dispersion relation of thin films of bulk hBN with film thicknesses of 1, 2, 4, and 8 times the interlayer spacing of bulk hBN, which is roughly \SI{0.33}{\nm}. For these plots, we take hBN to be cleaved such that the optical axis is perpendicular to the plane of the film. The components of the bulk permittivity perpendicular and parallel to the c-axis ($\epsilon_{\perp}$ and $\epsilon_{\parallel}$, respectively; indexed by $\alpha\in\{\perp,\parallel\}$ below) are then given by
\begin{equation}
	\epsilon_{\alpha}(\omega) = \epsilon_{\infty,\alpha}\Bigg(1 + \frac{\omega^2_{\mathrm{LO},\alpha}-\omega^2_{\mathrm{TO},\alpha}}{\omega^2_{\mathrm{TO},\alpha}-\omega^2} \Bigg),
	\label{eq:3deps}
\end{equation}
with $\epsilon_{\infty,\parallel} = \num{2.95}$, $\omega_{\mathrm{TO},\parallel} = \SI{760}{\per\cm}$, and $\omega_{\mathrm{LO},\parallel} = \SI{830}{\per\cm}$; and $\epsilon_{\infty,\perp} = \num{4.87}$, $\omega_{\mathrm{TO},\perp} = \SI{1360}{\per\cm}$, and $\omega_{\mathrm{LO},\perp} = \SI{1614}{\per\cm}$\cite{caldwell2014sub,dai2014tunable}.
Losses are ignored in this discussion altogether. 
In the range between $\omega_{\mathrm{TO}, \alpha}$ and $\omega_{\mathrm{LO}, \alpha}$, the corresponding component of the permittivity is negative, while the other component is positive. This hyperbolicity leads to a dispersion for hBN thin films that have multiple branches at a given frequency, as can be seen in Figure~\ref{fig:2}(c).  This trend persists even when the thickness of the bulk is taken down to a single layer, albeit being pushed to high wavevectors. This is in contrast with the true monolayer, in which there is only one LO phonon mode. Given that the phonon polariton of 2D is the LO phonon, there can be only one branch of the dispersion. The bulk dielectric function of Eq.~\eqref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity. For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode. However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Figure~\ref{fig:2}(c), because even for a polariton wavelength of \SI{12}{\nm}, which is very short, the polariton frequency is only a few \si{\per\cm} above the TO frequency, within the damping linewidth of the TO frequency. %\textcolor{blue}{The bulk dielectric function of Eq.~\ref{eq:3deps} fails to capture this lack of higher-order modes, and thus hyperbolicity. For a bilayer however, there is a higher-order LO mode, associated with out-of-phase oscillation between the two layers, and thus one expects to see a higher-order mode. However, it is expected challenging to observe such a mode, as can be anticipated from the blue higher-order mode in Figure~\ref{fig:2}(c), because even for a polariton wavelength of 12 nm, which is very short, the polariton frequency is only a few cm$^{-1}$ above the TO frequency, within the damping linewidth of the TO frequency.}

Another anomaly between 2D and 3D is that the dispersion relation of the 2D LO phonon is non-monotonic, and the frequency starts to decrease for sufficiently large wavevector. This can never happen by extrapolating to a monolayer a dielectric function which has finite LO--TO splitting. This turn-back point corresponds to a frequency of around \SI{1500}{\per\cm} and a phonon wavelength of less than \SI{1}{\nm}. Thus, this deviation is challenging to probe in typical optical experiments. However it means that for the monolayer, the range of accessible phonon polaritonic frequencies for optical applications is reduced compared to thick films. Finally, while we have focused on anomalies between the atomically-thin limit of bulk and a true monolayer, a comparison of the monolayer with the fundamental phonon polariton mode of the one-atom-thick thin film suggests that we have a reasonable qualitative understanding of the monolayer dispersion from this limit. This qualitative similarity is to be somewhat expected, as in many cases, a 2D layer can be considered as a very thin film of bulk, provided that the microscopic properties of the bulk and monolayer are the same. In fact, if we take a thin film of thickness $t$ of hBN of dielectric function given by Eq.~\eqref{eq:3deps}, and focus only on the in-plane response, then there is a mathematical condition for the thin film to exactly mirror the bulk, namely
\begin{equation}
(\omega^{2 (\mathrm{bulk})}_{\mathrm{LO}} - \omega^{2 (\mathrm{bulk})}_{\mathrm{TO}})t = 4\omega^{(\mathrm{mono})}_{\mathrm{TO}}v^{(\mathrm{mono})}_g,
\end{equation}
where `bulk' denotes a property of the bulk dielectric function and `mono' denotes that of the monolayer conductivity.
